---
title: "432 Homework 5"
output: github_document
---

# General Instructions

Deadline: 2019-04-12 at 2 PM. Submit via [Canvas](https://canvas.case.edu/). 

Your response should include an R Markdown file and an HTML document that is the result of applying your R Markdown file.

The data set for this assignment is the same `oh_counties_2017.csv` data set that you worked with in Homework 4.

Start a separate R Project for Homework 5, as your first step. 

# Question 1 (10 points)

I have provided three essay prompts below. Select **one** of the three available prompts and answer it. Start the essay by clearly labelling the prompt you are using. We are looking for well-written, clear and thoughtful responses. Going a little over the suggested length is OK. Under is not OK.

- a) In the section on scientific blogging in Jeff Leek's *How to Be A Modern Scientist*, he remarks that "Everyone is an internet scientist now." In 5-10 complete English sentences, I want you to specify, using your own words and complete English sentences, one or two different ways in which you anticipate your work changing in light of this remark. What might you do differently, if you thought of yourself as an internet scientist? Do you, already, feel this way? Does this make being a scientist seem more or less appealing to you, and why? Please feel encouraged to provide an example from your own experiences that helps you explain your reactions to this idea.

- b) In 5-10 complete English sentences, I want you to specify, using your own words and complete English sentences, the least helpful piece of advice you took away from reading Jeff Leek's *How To Be A Modern Scientist*. Please provide a reference to the section of the book that provides this problematic advice. Please avoid selecting something because it does not apply to your current situation, but instead identify something that might be relevant to you, but that you are actually unwilling to do, and explain why. Please feel encouraged to provide an example from your own experiences that helps you explain why you are unwilling to take this piece of advice.

- c) Reread the section of Jeff Leek's *How To Be A Modern Scientist* on data sharing, where, among other things, he states that "the reproducibility debate is over." In 5-10 complete English sentences, I want you to specify how the "debate" over reproducibility has affected your field, and what changes you have made (or intend to make) in light of what you've learned in this course, from Leek's book, and from your experiences over the last year to make your work more reproducible. How important is building reproducible science in your field? In your view, what does "reproducible science" mean to you, and to the people you will be working with as you complete your degree program? Please feel encouraged to provide an example from your own experiences that helps you explain your reactions to the notion of reproducible science.

## Data Set for this Assignment (repeated from Homework 4)

The `oh_counties_2017.csv` data set I have provided describes a series of variables, pulled from the data for the 88 counties of the the State of Ohio from the [County Health Rankings](http://www.countyhealthrankings.org/rankings/data/oh) report for 2017. 

- You may also be interested in looking at the details of the [2017 Ohio Summary report (pdf)](http://www.countyhealthrankings.org/sites/default/files/state/downloads/CHR2017_OH.pdf), or at the [Excel data file](http://www.countyhealthrankings.org/sites/default/files/state/downloads/2017%20County%20Health%20Rankings%20Ohio%20Data%20-%20v2.xls) from which I created the `oh_counties_2017.csv` file. 
- Note that the 2018 data and report were released recently, but we will concentrate in this assignment on the 2017 results.

The available variables are listed below.  Each variable describes data at the **COUNTY** level.

Variable | Description
:-----------: | ------------------------------------------------------------------------------
fips | Federal Information Processing Standard code
county | name of County
years_lost_rate | age-adjusted years of potential life lost rate (per 100,000 population)
sroh_fairpoor | % of adults reporting fair or poor health (via BRFSS)
phys_days | mean number of reported physically unhealthy days per month
ment_days | mean number of reported mentally unhealthy days per mo
lbw_pct | % of births with low birth weight (< 2500 grams)
smoker_pct | % of adults that report currently smoking
obese_pct | % of adults that report body mass index of 30 or higher
food_env | indicator of access to healthy foods, in points (0 is worst, 10 is best)
inactive_pct | % of adults that report no leisure-time physical activity
exer_access | % of the population with access to places for physical activity
exc_drink | % of adults that report excessive drinking
alc_drive | % of driving deaths with alcohol involvement
sti_rate | Chlamydia cases / Population x 100,000
teen_births | Teen births / females ages 15-19 x 1,000
uninsured | % of people under age 65 without insurance
pcp_ratio | Population to Primary Care Physicians ratio
prev_hosp | Discharges for Ambulatory Care Sensitive Conditions/Medicare Enrollees x 1,000
hsgrads | High School graduation rate
unemployed | % of population age 16+ who are unemployed and looking for work
poor_kids | % of children (under age 18) living in poverty
income_ratio | Ratio of household income at the 80th percentile to income at the 20th percentile
associations | # of social associations / population x 10,000
pm2.5 | Average daily amount of fine particulate matter in micrograms per cubic meter
h2oviol | Presence of a water violation: Yes or No
sev_housing | % of households with at least 1 of 4 housing problems: overcrowding, high housing costs, or lack of kitchen or plumbing facilities
drive_alone | % of workers who drive alone to work
age.adj.mortality | premature age-adjusted mortality
dm_prev | % with a diabetes diagnosis
freq_phys_distress | % in frequent physical distress
freq_mental_distress | % in frequent mental distress
food_insecure | % who are food insecure
insuff_sleep | % who get insufficient sleep
health_costs | estimated mean health care costs
median_income | estimated median income
population | population size 
age65plus | % of population who are 65 and over
african-am | % of population who are African-American
hispanic | % of population who are of Hispanic/Latino ethnicity
white | % of population who are White
female | % of population who are Female
rural | % of people in the county who live in rural areas 

## Note for Questions 2-4

For Questions 2-4, you're going to develop a series of models using **86** of the counties (every county other than Cuyahoga County and Monroe County). In each case, you will fit and select a model using that sample of 86 counties, and then be asked to use that model to make predictions of the outcome of interest for Cuyahoga County and for Monroe County and to assess the quality of those predictions.
 
# Question 2 (10 points)

Build a reasonable linear or generalized linear model in your development sample (86 counties) to predict one of the outcomes in the `oh_counties_2017.csv` data set that describes a percentage (that must fall between 0 and 100) effectively using at least three and no more than 6 other variables from the list above. Demonstrate how well the model fits as well as the conclusions you draw from the model carefully. Be sure to discuss model assumptions. Then use the model to predict Cuyahoga County and Monroe County results, and assess the quality of those predictions.

# Question 3 (15 points)

Divide the 86 counties in your development sample into three groups (low, middle and high) in a rational way in terms of the `years_lost_rate` outcome. Make that new grouping your outcome for an ordinal logistic regression model. Fit a model (using a carefully selected group of no more than 5 predictor variables) and assess its performance carefully. Do not include the `age65plus` variable as a predictor, as the `years_lost_rate` data are age-adjusted already. Demonstrate how well the model fits as well as the conclusions you draw from the model carefully. Then use the model to predict Cuyahoga County and Monroe County results, and assess the quality of those predictions.

## A Hint (for Questions 3 and 4, in particular)

`polr` and several of the other modeling approaches we've worked on recently are finicky, at least in comparison to OLS. Sometimes, you'll get to the point where it seems like the model won't run, or won't summarize properly, or you have some extremely large or extremely small coefficient estimates or standard errors. Should this happen to you, the first thing I would do is try to identify which of your predictors is causing this problem, by running the model first with one predictor, then two, etc. until you figure out which predictors cause problems. Reasons why you could be having a problem include:

1. a predictor has values that completely identify the category of your outcome variable, perfectly (e.g., one category's predictor values are inevitably lower than all of another category's predictor values, with no overlap)
2. the scales of the predictors are wildly different, for instance one predictor has extremely large or extremely small values, causing the estimated standard errors to explode, which should cause you to think about reducing the impact of that, perhaps by changing the units, say from $s to $1000s or by standardizing the predictors
3. intense collinearity between two or more of your predictors
4. coding issues in setting up one or more of the variables.

For example, some people in the past have tried to use `median_income` in their models along with other variables that have much smaller scales (ranges). I would try rescaling those predictors with large ranges to have similar magnitudes to the other predictors, perhaps simply by expressing the median income in thousands of dollars (by dividing the raw data by 1000) rather than on its original scale, or perhaps by standardizing all of the coefficients by subtracting their means and dividing by their standard deviations.

As another example, some people in the past tried using age-adjusted mortality to predict years lost rate, but if you divide the years lost rate into several ordinal categories, it's not hard to wind up in a situation where age-adjusted mortality is perfectly separated, so that if you know the mortality, it automatically specifies the years lost rate category in these data.

# Question 4 (15 points)

Build a new outcome variable that is a count (possible values = 0-4) of whether the county meets each of the following standards:

- the county has a `smoker_pct` value **below** the Ohio-wide average of 22
- the county has an `obese_pct` value **below** the Ohio-wide average of 31
- the county has an `exer_access` value **above** the Ohio-wide average of 83
- the county has **NOT** had a water violation in the past year (as shown by `h2oviol` = No)

To illustrate, consider these five counties:

County | `smoker_pct` | `obese_pct` | `exer_access` | `h2oviol` | Standards Met
-------------: | --------: | --------: | --------: | --------: | --------:
*Standard* | < 22 | < 31 | > 83 | No | --
Highland | 23.73 | 32.1 | 57.66 | Yes | 0
Erie | **18.99** | 35.4 | 80.04 | Yes | 1
Ashland | **19.79** | **27.9** | 64.65 | Yes | 2
Athens | 23.34 | **28.3** | **85.00** | **No** | 3
Cuyahoga | **18.70** | **30.0** | **95.55** | **No** | 4

Your job is to fit **two** possible regression models in your development sample to predict this count, using the same predictors (at least 3 and no more than 6 of those not used in the calculation of standards) available in the data set. Demonstrate how well each model fits the counts by developing a rootogram and other summaries that you deem useful, then select the model you prefer, specifying your reasons. Next, use your preferred model to predict Cuyahoga County and Monroe County results, and assess the quality of those predictions.

## Please add the session information.

Finally, at the end of this homework and all subsequent assignments (including the projects), please add the session information. My preferred way for you to do  this is:

```{r, eval = FALSE}
sessioninfo::session_info()
```

which will produce something like this:

```{r}
sessioninfo::session_info()
```